Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-10-03 prop_letter.tex
prop_letter.tex \section{Gleason} Updating...

In Sec.~\ref{sec:Interval-Uncertainty}, we discussed how states
consistent with a QIVPM~$\bar{\mu}$ on the projectors in a subspace
of event~$\events'$. In particular, if $\events'$ commutes, $\bar{\mu}$
restricting on $\events'$ is essentially a classical IVPM, and Thm.~\ref{thm:Shapley}
claims that there is always a state~$\rho$ consistent with $\bar{\mu}$.
In this section, we want to shift our attention to consider states
consistent with $\bar{\mu}$ on every event, $\coreBorn\left(\bar{\mu},\events\right)$,
so that these states could characterize $\bar{\mu}$ comprehensively.
When the intervals are infinitely precise, i.e., $\mathscr{I}_{\infty}=\set{\left[x,x\right]}{x\in\left[0,1\right]}$,
a QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{\infty}$ becomes
quantum probability measure~\citep{HOSTunpublished}. In this situation,
Gleason's theorem claims that there is always a unique state consistent
with $\bar{\mu}$~\citep{gleason1957,Redhead1987-REDINA,peres1995quantum},
and this state can easily induce $\bar{\mu}$ according to the Born
rule~\citep{Born1983,peres1995quantum,544199,Jaeger2007}. However,
if we consider general QIVPMs mapping to an \emph{arbitrary} set of
intervals, there is a QIVPM that no state can be consistent with it.

@@

In other words, if the measurement intervals are too imprecise, the
results we get might contradict to each other, and exclude any possible
state from being consistent with the measurement results. We will
address this problem with a concrete family of intervals and QIVPMs,
and leave for another time the general framework~\citep{HOSTunpublished}.
Consider the following intervals 
\begin{equation}
\begin{aligned}\mathscr{I}_{n}= & \set{\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]\cap\left[0,1\right]}{k=0,\ldots,2n}\\
 & \cup\{\imposs,\necess\}
\end{aligned}
\end{equation}
which are overlapping intervals with length~$1/n$ except for the
intervals with length~$\left(2n\right)^{-1}$ at each end. As shown
in Table~\ref{tab:I_n}, for each interval~$[\ell,r]$, we define
$\textrm{Int}\left([\ell,r]\right)$, the non-overlapping interior
part of $[\ell,r]$.
\begin{table}
\caption{\label{tab:I_n}Intervals in $\mathscr{I}_{n}$, their lengths, and
their interiors, where $2\mid k$ means $k$ is even, and $2\nmid k$
means $k$ is odd. Although intervals~$[\ell,r]$ are overlapping,
their interior~$\textrm{Int}\left([\ell,r]\right)$ are non-overlapping.}

@@

To prove that $\left[\mu_{\rho}^{\mathrm{B}}\right]_{n}$ is a QIVPM,
it is sufficient to check Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}).
Given $x_{0}$, $x_{1}$, and $x_{0}+x_{1}\in\left[0,1\right]$,
\begin{equation}
\iota_{n}\left(x_{0}+x_{1}\right)\subseteq\iota_{n}\left(x_{0}\right)+\iota_{n}\left(x_{1}\right)\label{eq:iota-add}
\end{equation}
implies Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
when $P_{0}$ and $P_{1}$ are orthogonal. When $P_{0}$ and $P_{1}$
are not orthogonal on a three-dimensional Hilbert space, we have two
possibilities:
\begin{itemize}
\item Their complements are orthogonal so that we can apply Eq.~(\ref{eq:iota-add})
on their complements.
\item One of them projects onto a subspace of what the other one projects
onto.
\end{itemize}
\noindent Both cases deduce Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
easily~\citep{HOSTunpublished} so that $\left[\mu_{\rho}^{\mathrm{B}}\right]_{n}$
is indeed a QIVPM.

>>>>

In Sec.~\ref{sec:Interval-Uncertainty}, we discussed how states
consistent with a QIVPM~$\bar{\mu}$ on the projectors in a subspace
of event~$\events'$. In particular, if $\events'$ commutes, $\bar{\mu}$
restricting on $\events'$ is essentially a classical IVPM, and Thm.~\ref{thm:Shapley}
claims that there is always a state~$\rho$ consistent with $\bar{\mu}$.
In this section, we want to shift our attention to consider states
consistent with $\bar{\mu}$ on \emph{every} event, $\coreBorn\left(\bar{\mu},\events\right)$,
so that these states could characterize $\bar{\mu}$ comprehensively.
When the intervals are infinitely precise, i.e., $\mathscr{I}_{\infty}=\set{\left[x,x\right]}{x\in\left[0,1\right]}$,
a QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{\infty}$ becomes
quantum probability measure~\citep{HOSTunpublished}. In this situation,
Gleason's theorem claims that there is always a unique state consistent
with $\bar{\mu}$~\citep{gleason1957,Redhead1987-REDINA,peres1995quantum},
and this state can easily induce $\bar{\mu}$ according to the Born
rule~\citep{Born1983,peres1995quantum,544199,Jaeger2007}. However,
if we consider general QIVPMs mapping to an \emph{arbitrary} set of
intervals, there is a QIVPM that no state can be consistent with it.

@@

In other words, if the measurement intervals are too imprecise, the
results we get might contradict to each other, and exclude any possible
state from being consistent with the measurement results. We will
address this problem with a concrete family of intervals and QIVPMs,
and leave for another time the general framework~\citep{HOSTunpublished}.
Consider the following intervals 
\begin{equation}
\begin{aligned}\mathscr{I}_{n}= & \set{\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]\cap\left[0,1\right]}{k=0,\ldots,2n}\\
 & \cup\{\imposs,\necess\}
\end{aligned}
\end{equation}
which are overlapping intervals with length~$1/n$ except for the
intervals with length~$\left(2n\right)^{-1}$ at each end. As shown
in Table~\ref{tab:I_n}, for each interval~$[\ell,r]$, we define
$\textrm{Int}\left([\ell,r]\right)$, the non-overlapping interior
part of $[\ell,r]$.
\begin{table}
\caption{\label{tab:I_n}Intervals in $\mathscr{I}_{n}$, their lengths, and
their interiors, where $2\mid k$ means $k$ is even, and $2\nmid k$
means $k$ is odd. Although intervals~$[\ell,r]$ are overlapping,
their interiors~$\textrm{Int}\left([\ell,r]\right)$ are non-overlapping.}

@@

To prove that $\left[\mu_{\rho}^{\mathrm{B}}\right]_{n}$ is a QIVPM,
it is sufficient to check Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}).
Given $x_{0}$, $x_{1}$, and $x_{0}+x_{1}\in\left[0,1\right]$,
\begin{equation}
\iota_{n}\left(x_{0}+x_{1}\right)\subseteq\iota_{n}\left(x_{0}\right)+\iota_{n}\left(x_{1}\right)\label{eq:iota-add}
\end{equation}
implies Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
when $P_{0}$ and $P_{1}$ are orthogonal. When $P_{0}$ and $P_{1}$
are not orthogonal on a three-dimensional Hilbert space, we have two
possibilities:
\begin{itemize}
\item Their complements~$\mathbb{1}-P_{0}$ and $\mathbb{1}-P_{1}$ are
orthogonal so that we can apply Eq.~(\ref{eq:iota-add}) on their
complements.
\item One of them projects onto a subspace of what the other one projects
onto.
\end{itemize}
\noindent Both cases deduce Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
easily~\citep{HOSTunpublished} so that $\left[\mu_{\rho}^{\mathrm{B}}\right]_{n}$
is indeed a QIVPM.
